---
title: "Causer Machine Learning Assignment"
author: "Javier Rios Perez"
date: "12/17/2016"
output: html_document
---

Coursera: Machine Learning - Assignment solution

Loading the required R libraries 
```{r}
library(ggplot2)
library(lattice)
library(caret)
library(RCurl)
```

Cleaning the data
Here I will remove the empty, zero and NA columns becuase they are not useful
for my analysis, and remove the all the factor columns but the classe variable.
This is done in the training set.

```{r echo = FALSE}
#reading the training data
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")

############################################################
# cleaning the data
# subseting the training test to get: 
# sub_training and sub_testing sets
############################################################

# cleaning the data - eliminating NA and zero colums as well as the factor cols that
# my model is not interested in

NA_cols <-  sapply(training, function(x){!any(is.na(x))})
clean_training <- training[,NA_cols]

ZERO_cols <- sapply(clean_training, function(x) {!any(x== "")})
clean_training <- clean_training[,ZERO_cols]
str(clean_training)

#remvoong times and other factos
clean_training <- clean_training[,-c(1:7)]

```

Creating a training teting sets from the original training set.

```{r include = FALSE}
# partitioning the clean_training database into 2 different ones
# sub_training and sub_testing

set.seed(1234)
inTrain <- createDataPartition(y = clean_training$classe, p = 0.75, list = FALSE)

sub_training <- clean_training[inTrain,]
sub_testing <- clean_training[-inTrain,]

```

Building the rpart model

```{r include = FALSE}

modFit_rpart <- train(classe ~., method = "rpart", data = sub_training)
predict_rpart <- predict(modFit_rpart, newdata = sub_testing)
confMatrix_rpart <- confusionMatrix(predict_rpart, sub_testing$classe)


```

```{r}
#rpart
print(confMatrix_rpart)
confMatrix_rpart_table <- as.matrix(confMatrix_rpart$table)
heatmap(as.matrix(confMatrix_rpart$table),
        scale = "column",
        color = heat.colors(256),
        Rowv = NA,
        Colv = NA,
        main = "Confusion Matrix - rpart method")
```


Building the rf model
```{r include = FALSE}
# rf

  modFit_rf <- train(classe ~., method = "rf", data = sub_training)
  predict_rf <- predict(modFit_rf, newdata = sub_testing)
  confMatrix_rf <- confusionMatrix(predict_rf, sub_testing$classe)

  
```

```{r} 
  print(confMatrix_rf)
  confMatrix_rf_table <- as.matrix(confMatrix_rf$table)
  heatmap(as.matrix(confMatrix_rf$table),
          scale = "column",
          color = heat.colors(256),
          Rowv = NA,
          Colv = NA,
          main = "Confusion Matrix - rf method")
```

Comparing results

```{r}
############################################################
# Results
############################################################

par(mfrow = c(1,2))
plot(confMatrix_rpart_table, main = "Confusion Matrix - rpart method")
plot(confMatrix_rf_table, main = "Confusion Matrix - rf method")  
  
```

